---
title: "Final Project DCS 375"
author: "Dieter Villegas"
date: "4/6/2022"
output: html_document
---


```{r load-libraries}
library(dplyr)
library(igraph)
```


```{r clean-data}

# read in the RDS
wc_rds <- readRDS("WC_events_clean.RDS")

# grab only the columns we're interested in for analysis
wc_maxtrix <- cbind(wc_rds$period,wc_rds$timestamp, wc_rds$duration,wc_rds$possession_team.name,wc_rds$type.name,wc_rds$player.name,wc_rds$team.name,wc_rds$pass.length,wc_rds$pass.recipient.name,wc_rds$pass.outcome.name,wc_rds$ball_receipt.outcome.name,wc_rds$match_id,wc_rds$location.x,wc_rds$location.y,wc_rds$pass.end_location.x,wc_rds$pass.end_location.y)

# turn into a dataframe
wc_df_official <- as.data.frame(wc_maxtrix)

# rename the columns for analysis
colnames(wc_df_official) <- c("period","timestamp","duration","possession_team.name","type.name","player.name","team.name","pass.length","pass.recipient.name","pass.outcome.name","ball_receipt.outcome.name","match_id","location.x","location.y","pass.end_location.x","pass.end_location.y")

# grab only the passes
network_passes <- subset(wc_df_official, type.name == "Pass")

# create 4 data sets for four games we plan to analyze
france_croatia_final <- subset(network_passes, possession_team.name == c("France", "Croatia"))

france_belgium_semi <- subset(network_passes, team.name == c("France", "Belgium"))

france_uruguay_quarter <- subset(network_passes, team.name == c("France", "Uruguay"))

france_argentina_round <- subset(network_passes, team.name == c("France", "Argentina"))

# write 4 csv files for each of these games

#write.csv(france_croatia_final, "france_croatia_final.csv", row.names = FALSE)

#write.csv(france_belgium_semi, "france_belgium_semi.csv", row.names = FALSE)

#write.csv(france_uruguay_quarter, "france_uruguay_quarter.csv", row.names = FALSE)

#write.csv(france_argentina_round, "france_argentina_round.csv", row.names = FALSE)

```


```{r france-final-object}


# read in csv so that we don't have to re-run top part in case of a premature change to players.
france_belgium_semi <- read.csv("france_belgium_semi.csv")

# clean the data frame to only include passes from france and croatia final
team_name_france_belgium_semi <- subset(france_belgium_semi, team.name == c("France", "Belgium"))

france_only_semi <- subset(team_name_france_belgium_semi, team.name == "France")
france_only_semi <- subset(france_only_semi, player.name != "Steve Mandanda")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Steve Mandanda")

france_only_semi <- subset(france_only_semi, player.name != "Thomas Lemar")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Thomas Lemar")

france_only_semi <- subset(france_only_semi, player.name != "Steven N''Kemboanza Mike Christopher Nzonzi")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Steven N''Kemboanza Mike Christopher Nzonzi")

france_only_semi <- subset(france_only_semi, player.name != "Presnel Kimpembe")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Presnel Kimpembe")

france_only_semi <- subset(france_only_semi, player.name != "Ousmane Dembélé")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Ousmane Dembélé")

france_only_semi <- subset(france_only_semi, player.name != "Nabil Fekir")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Nabil Fekir")

france_only_semi <- subset(france_only_semi, player.name != "Florian Thauvin")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Florian Thauvin")

france_only_semi <- subset(france_only_semi, player.name != "Benjamin Mendy")
france_only_semi <- subset(france_only_semi, pass.recipient.name != "Benjamin Mendy")



belgium_only_semi <- subset(team_name_france_belgium_semi, team.name == "Belgium")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Thomas Vermaelen")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Thomas Vermaelen")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Dries Mertens")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Dries Mertens")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Thorgan Hazard")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Thorgan Hazard")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Youri Tielemans")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Youri Tielemans")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Adnan Januzaj")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Adnan Januzaj")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Anga Dedryck Boyata")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Anga Dedryck Boyata")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Leander Dendoncker")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Leander Dendoncker")




# construct dataset for connected passes only 
connected_pass_france_semi <- subset(france_only_semi, is.na(pass.outcome.name))
connected_pass_belgium_semi <- subset(belgium_only_semi, is.na(pass.outcome.name))

# create an edgelist of passes to and from players
france_semi <- cbind(connected_pass_france_semi$player.name, connected_pass_france_semi$pass.recipient.name)

belgium_semi <- cbind(connected_pass_belgium_semi$player.name, connected_pass_belgium_semi$pass.recipient.name)

# name these columns
colnames(france_semi) <- c("from", "to")
colnames(belgium_semi) <- c("from", "to")

# create a graph object for network analysis
france_semi.g <- graph.edgelist(france_semi, directed = TRUE) # makes a graph object
belgium_semi.g <- graph.edgelist(belgium_semi, directed = TRUE) # makes a graph object

png(file="france_semi.png", width=1000, height=1000)
plot(france_semi.g, )
png(file="belgium_semi.png", width=1000, height=1000)
plot(belgium_semi.g)
dev.off()
```

```{r identify-france-players}

france_midfielder <- c("Corentin Tolisso", "Paul Pogba", "N''Golo Kanté", "Blaise Matuidi", "Steven N''Kemboanza Mike Christopher Nzonzi", "Thomas Lemar", "Florian Thauvin")
france_defender <- c("Benjamin Pavard", "Samuel Yves Umtiti", "Raphaël Varane", "Lucas Hernández Pi", "Presnel Kimpembe", "Djibril Sidibé", "Benjamin Mendy")
france_forward <- c("Ousmane Dembélé", "Kylian Mbappé Lottin", "Antoine Griezmann", "Nabil Fekir", "Olivier Giroud")
france_goalkeeper <- c("Hugo Lloris", "Steve Mandanda")
```

```{r identify-belgium-players}

belgium_midfielder <- c("Kevin De Bruyne", "Axel Witsel", "Yannick Ferreira Carrasco", "Mousa Sidi Yaya Dembélé", "Thorgan Hazard", "Nacer Chadli", "Marouane Fellaini-Bakkioui", "Youri Tielemans", "Adnan Januzaj")
belgium_defender <- c("Thomas Meunier", "Anga Dedryck Boyata", "Jan Vertonghen", "Toby Alderweireld", "Leander Dendoncker", "Thomas Vermaelen", "Vincent Kompany")
belgium_forward <- c("Eden Hazard", "Dries Mertens", "Romelu Lukaku Menama", "Michy Batshuayi Tunga")
belgium_goalkeeper <- c("Thibaut Courtois")
```

```{r}
V(france_semi.g)[V(france_semi.g)$name %in% france_midfielder]$color <- "#440154"
V(france_semi.g)[V(france_semi.g)$name %in% france_defender]$color <- "#31688e"
V(france_semi.g)[V(france_semi.g)$name %in% france_forward]$color <- "#35b779"    
V(france_semi.g)[V(france_semi.g)$name %in% france_goalkeeper]$color <- "#fde725"
```
#### Do centrality and betweenesss 
- take out players that did not actually play 

```{r}
png(file="france_semi.png", width=1000, height=1000)
par(bg="transparent")
plot(france_semi.g, vertex.size = 20, vertex.label.cex = 1, vertex.label.color = "white")
dev.off()


```

# adding the soccer field as a background to the networks 
```{r}
#install.packages("magick")
library(magick)
```

```{r}

soccer_png <- image_read("soccer_fil.png")
france_png <- image_read("france_semi.png")


```


```{r}
# so we dont have to re-read images from above
france_pgn1 <- france_png
soccer_field <-soccer_png
# scaling the soccer_field to 1000 in height and having it without a background 
#soccer_field <- image_scale(image_background(soccer_field, "none"), "x1000")

#scaling france network of passes image to 1000 height
france_pgn1 <- image_scale(france_pgn1, "x1000")
#creating a composite of both images and moving the France network using offset to the right and down
# scaling the soccer field to 1000 in height 
final_image <- image_composite(image_scale(soccer_field, "x1000"), france_pgn1, offset = "+500+50")

```

```{r}
#exporting image to folder 
image_write(final_image,path = "semis_field_france.png", format = "png")
```

website explaning how to use magick 
https://cran.r-project.org/web/packages/magick/vignettes/intro.html



maybe isntead of weighted edges we switch node size depending on the amount of connections they have?
but if we can fiugre out edge weight would be hugeee
```{r}

#png(file="france_semi.png", width=1000, height=1000)
#par(bg="transparent")
deg <- degree(france_semi.g, mode="all")
#plot(france_semi.g, vertex.size = deg/4, vertex.label.cex = 1, vertex.label.color = "white")
#dev.off()



```







```{r}
trial_france <-france_semi

trial_edlist_france<- get.edgelist(trial_france)


#cbind( get.edgelist(g) , round( E(g)$weight, 3 ))
```

```{r}

ee <- c(1,2,2,3,1,2,2,4,1,2,2,3)
g <- graph(trial_france)

plot(g)

# Returns a two column matrix with edge names and counts
edge.count1 <- function(g){
    result <- cbind(E(g), count.multiple(g))
    colnames(result) <- c("edge", "count")
    result
}
# Returns a three column matrix with edges and counts
edge.count2 <- function(g){
  cbind(get.edgelist(g), count.multiple(g))
}
# Returns a data.frame of unique edges and counts
edge.count3 <- function(g){
     D <- data.frame(get.edgelist(g))  # convert to data frame
     ones <- rep(1, nrow(D))   # a column of 1s
     result <- aggregate(ones, by = as.list(D), FUN = sum)
     names(result) <- c("from", "to", "count")
     result
}
edge.count3(g)


```
```{r}

```



run a for loop for histogram 


how france plays, 



 