---
title: "new"
author: "Jenniflore Beaubrun"
date: '2022-04-19'
output: html_document
---

```{r load-libraries}
library(dplyr)
library(igraph)
```


```{r clean-data}

# read in the RDS
wc_rds <- readRDS("WC_events_clean.RDS")

# grab only the columns we're interested in for analysis
wc_maxtrix <- cbind(wc_rds$period,wc_rds$timestamp, wc_rds$duration,wc_rds$possession_team.name,wc_rds$type.name,wc_rds$player.name,wc_rds$team.name,wc_rds$pass.length,wc_rds$pass.recipient.name,wc_rds$pass.outcome.name,wc_rds$ball_receipt.outcome.name,wc_rds$match_id,wc_rds$location.x,wc_rds$location.y,wc_rds$pass.end_location.x,wc_rds$pass.end_location.y)

# turn into a dataframe
wc_df_official <- as.data.frame(wc_maxtrix)

# rename the columns for analysis
colnames(wc_df_official) <- c("period","timestamp","duration","possession_team.name","type.name","player.name","team.name","pass.length","pass.recipient.name","pass.outcome.name","ball_receipt.outcome.name","match_id","location.x","location.y","pass.end_location.x","pass.end_location.y")

# grab only the passes
network_passes <- subset(wc_df_official, type.name == "Pass")

# create 4 data sets for four games we plan to analyze
france_croatia_final <- subset(network_passes, possession_team.name == c("France", "Croatia"))

france_belgium_semi <- subset(network_passes, team.name == c("France", "Belgium"))

france_uruguay_quarter <- subset(network_passes, team.name == c("France", "Uruguay"))

france_argentina_round <- subset(network_passes, team.name == c("France", "Argentina"))

# write 4 csv files for each of these games

#write.csv(france_croatia_final, "france_croatia_final.csv", row.names = FALSE)

#write.csv(france_belgium_semi, "france_belgium_semi.csv", row.names = FALSE)

#write.csv(france_uruguay_quarter, "france_uruguay_quarter.csv", row.names = FALSE)

#write.csv(france_argentina_round, "france_argentina_round.csv", row.names = FALSE)

```


```{r belgium-final-object}


# read in csv so that we don't have to re-run top part in case of a premature change to players.
france_belgium_semi <- read.csv("france_belgium_semi.csv")

# clean the data frame to only include passes from france and croatia final
team_name_france_belgium_semi <- subset(france_belgium_semi, team.name == c("France", "Belgium"))

belgium_only_semi <- subset(team_name_france_belgium_semi, team.name == "France")


belgium_only_semi <- subset(team_name_france_belgium_semi, team.name == "Belgium")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Thomas Vermaelen")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Thomas Vermaelen")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Dries Mertens")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Dries Mertens")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Thorgan Hazard")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Thorgan Hazard")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Youri Tielemans")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Youri Tielemans")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Adnan Januzaj")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Adnan Januzaj")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Anga Dedryck Boyata")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Anga Dedryck Boyata")

belgium_only_semi <- subset(belgium_only_semi, player.name != "Leander Dendoncker")
belgium_only_semi <- subset(belgium_only_semi, pass.recipient.name != "Leander Dendoncker")



#construct dataset for connected passes only 
connected_pass_belgium_semi <- subset(belgium_only_semi, is.na(pass.outcome.name))

# create an edgelist of passes to and from players

belgium_semi <- cbind(connected_pass_belgium_semi$player.name, connected_pass_belgium_semi$pass.recipient.name)

# name these columns
colnames(france_semi) <- c("from", "to")
colnames(belgium_semi) <- c("from", "to")

# create a graph object for network analysis
belgium_semi.g <- graph.edgelist(belgium_semi, directed = TRUE) # makes a graph object

png(file="belgium_semi.png", width=1000, height=1000)
plot(belgium_semi.g)
dev.off()
```

```{r identify-belgium-players}

belgium_midfielder <- c("Kevin De Bruyne", "Axel Witsel", "Yannick Ferreira Carrasco", "Mousa Sidi Yaya Dembélé", "Thorgan Hazard", "Nacer Chadli", "Marouane Fellaini-Bakkioui", "Youri Tielemans", "Adnan Januzaj")
belgium_defender <- c("Thomas Meunier", "Anga Dedryck Boyata", "Jan Vertonghen", "Toby Alderweireld", "Leander Dendoncker", "Thomas Vermaelen", "Vincent Kompany")
belgium_forward <- c("Eden Hazard", "Dries Mertens", "Romelu Lukaku Menama", "Michy Batshuayi Tunga")
belgium_goalkeeper <- c("Thibaut Courtois")
```

```{r}
V(belgium_semi.g)[V(belgium_semi.g)$name %in% belgium_midfielder]$color <- "#440154"
V(belgium_semi.g)[V(belgium_semi.g)$name %in% belgium_defender]$color <- "#31688e"
V(belgium_semi.g)[V(belgium_semi.g)$name %in% belgium_forward]$color <- "#35b779"    
V(belgium_semi.g)[V(belgium_semi.g)$name %in% belgium_goalkeeper]$color <- "#fde725"
```
#### Do centrality and betweenesss 
- take out players that did not actually play 

```{r}
plot(belgium_semi.g, vertex.size = 20, vertex.label.cex = 0.5, vertex.label.color = "black")
```

#### Betweeenness
Run the following code to calculate betweenness scores for the top ten most between people in this network.  

Betweenness centrality is a way of detecting the amount of influence a node has over the flow of information in a graph

```{r}
btwn.team <- betweenness(belgium_semi.g)
names(btwn.team) <- V(belgium_semi.g)$name
ind <- order(-btwn.team)
btwn.team[ind][1:10]
``` 
#### Centrality
#Most Central Players 
Centrality scores for the top ten most central Player in this network

```{r}
cent.eig <- evcent(belgium_semi.g)
names(cent.eig$vector) <- V(belgium_semi.g)$name
ind <- order(-cent.eig$vector)
cent.eig$vector[ind][1:10]
```
#Least Central Players 
```{r}
cent.eig <- evcent(belgium_semi.g)
names(cent.eig$vector) <- V(belgium_semi.g)$name
ind <- order(-cent.eig$vector)
cent.eig$vector[ind][8:13]
```

#Histogram for Belgium Players 
```{r}

```






